---
title: "Workflow with quantatools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow with quantatools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7
)
```

```{r setup}
library(quantatools)
library(dplyr)
library(ggplot2)
```

*quantatools* is an implementation for Kendall's quanta hunting algorithm [@kendallHuntingQuanta1974] in R. It is used to identify e.g. the base measurement that was used when generating a sequence of lengths, which can be helpful for archaeologists and architects working with building measurements or volumes looking to find the unit-variation that was used in the making of artifact groups. The package es suited to gather any base measurement of a number of measurements and is not restricted to these use cases.

# Example

To demonstrate the packages functionality, we create a data set that has an approximate base measurement of 0.37:

```{r}
data <- 0.37 * round(rnorm(n = 100, mean = 100, sd = 50), digits = 0)
```

The data may be 'unrounded' before further processing to create a third decimal number and add some noise that will make the result more resistant to actual measurement errors or previous rounding errors [@kubickaMetrologicalResearchMachu2020, 90-91].

```{r}
data 

unround_data(data)
```

The function `calculate_quantogram()` employs the unrounding automatically. It can be deactivated when setting "unround" to FALSE. The function generates a data frame in which each quantum in the column `q` is given a value (f(q) in column `f_q`) representing its probability to be the base measurement.

```{r}
calculate_quantogram(data, unround = TRUE) %>% 
  head()
```

*quantatools* includes a plot method for this quantogram which produces plots such as the following:

```{r}
calculate_quantogram(data) %>% 
  plot_quantogram(title = "A Quantogram Plot")
```

The second peak that can be observed should be a 0.185, as the plotting function informs via message, as it is 0.37 / 2. The quantum at which f(q) peaks is likely to be interpreted as the base measure of the original data. As we stated before, we created data that has a base measurement of 0.37, which is the result returned by this quantogram.

# Erechtheion ("Real World Data")

To test this with real data, we added Table 1 ("Erechtheion, Athens. Building dimension.") from @pakkanenDerivingAncientFoot2002 :

```{r}
data("Erechtheion")
Erechtheion
```

Calculating and plotting the quantogram for these measurements suggests that the underlying base measure should be about 0.325m:

```{r echo = FALSE}
Erechtheion_quant <- calculate_quantogram(Erechtheion$Measurement_m)
```

```{r}
calculate_quantogram(Erechtheion$Measurement_m) %>%
  plot_quantogram(title = "A Quantogram Plot of Measurements from the Erechtheion")
```

This roughly corresponds with the foot measure that could be estimated by dividing the actual measurements by the number of feet they should encompass according to the buildings inscriptions:

```{r}
Erechtheion$Measurement_m / Erechtheion$FeetAfterInscr
```

The mean of all those measurements is `r round(mean(Erechtheion$Measurement_m / Erechtheion$FeetAfterInscr), digits = 3)`, which is very close to our calculated quantum of `r Erechtheion_quant$q[which.max(Erechtheion_quant$f_q)]`.

# Bootstrapping

With `bootstrap_quanta()` *quantatools* provides a way of determining the quantum using repeated resampling from the original data. The histogram will show the most probable quantum and its confidence intervals [@kubickaMetrologicalResearchMachu2020, 91].

The results can be easily plotted with `plot_bootstrap()`, which expects an object as returned by `bootstrap_quanta()`

```{r}
bootstr <- bootstrap_quanta(Erechtheion$Measurement_m, n_samples = 1000)

plot_bootstrap(bootstr)
```

The list returned by `bootstrap_quanta()` will also show the quanta, individual results and confidence intervals:

```{r}
bootstr$quanta
head(bootstr$results)
bootstr$`confidence interval`
```

# Measurement Errors

In order to understand how good the quantum works on the data, *quantatools* provides a function which calculates the difference between the 'ideal' measurement and the real ones. These can be displayed as a vector of single deviations (`mean = FALSE`, default) or as the mean error across all measurements (`mean = TRUE`).

```{r}
calculate_error(input = Erechtheion$Measurement_m, 
                quantum = get_quantum(Erechtheion$Measurement_m))

calculate_error(input = Erechtheion$Measurement_m, 
                quantum = get_quantum(Erechtheion$Measurement_m),
                mean = TRUE)
```
